---
title: "Intrepid-2a replication analyses"
author: "Marius 't Hart"
date: "2025-04-11"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview

The Intrepid consortium aims to compare three computational theories of consciousness by testing their predictions on spatial perception in inactivated areas of cortex. In the Intrepid-2a project, there are 3 psychophysics experiments testing spatial perception around the blind spot. In Toronto we are replicating the work from the primary lab in Glasgow. This consists of three psychophysics tasks that we analyse here.

An overview of the full Intrepid project can be found here: [https://osf.io/4rn85](https://osf.io/4rn85).

Tasks:

- area perception: matching task; a circle around the fovea has to matched in size to a peripheral circle which is either around the blind spot, or at a different area of the retina
- curvature perception: staircase task; participants reply whether an apparent motion path is curved to the left or to the right, the paths are either right next to the blind spot or away from the blind spot
- distance perception: staircase task; participants reply which of two dot pairs is further from each other, either one of the dot pairs spans the blind spot, or neither do

# Setup Code

First we load all the scripts to do the analyses:

```{r}
# this script has function to download the data:
source("R/data.R")
# this script contains generic utility functions:
source("R/utilities.R")
# script with custom psychometric functions:
source("R/mprobit.R")

# this script contains functions to process the distance task data:
source('R/distance2.R')
source('R/distance3.R') # checks on original tests
# this script contains functions to process the curvature task data:
source('R/curvature.R')
# this script contains functions to process the area task data:
source("R/area.R")
# alternative just using a gaussian density distribution:
source("R/area2.R")

# scripts testing relations between effects in distance and curvature task:
source("R/dist_curve.R")
```

# Download Data

We download the data from a (private) OSF repository:

```{r eval=F}
getData()
```

# Area task

A first peek at the data, plotted as an empirical cumulative distribution, fitted with a probit function:

```{r fig.width=6.5, fig.height=5}
plotAreaPsychometricCurve()
```

And here is an alternative, just plotting a density distribution:

```{r fig.width=6.5, fig.height=5}
plotAreaDensityCurve()
```

Statistics. Not expecting any effects.

We fit a probit function to each participants data in the 4 conditions, to determine the PSE/mean as well as the slope/sd, and can then do two repeated-measures ANOVA using eye (ipsi/contra) and location (across/away) as factors. One ANOVA on the PSE, and one on the slope.


```{r}
doAreaStats()
```

Indeed nothing is significant.

# Curvature task

Let's look at the data:

```{r fig.width=6.5, fig.height=5}
plotCurvaturePsychometricCurve()
```

Perhaps there is a small effect of whether the stimulus was presented close to the blind spot or away from it. We repeat the same ANOVA procedure as above.

```{r}
doCurvatureStats()
```


# Distance task

First, we look at the data:

```{r fig.width=6.5, fig.height=5}
plotDistancePsychometricCurve()
```

We can see differences in the bias, and perhaps in the slope. To test this statistically, we will again fit probit functions to each participants' data (all 4 conditions), and use the factors eye (ipsi/contra) and location (across/away) in two repeated measures ANOVA's: one for each parameter (bias and slope).

```{r}
doDistanceStats()
```

Let's plot the individual participant's PSEs and slopes in each condition to visually inspect for normality and outliers:

```{r fig.width=8, fig.height=5}
plotDistancePSEslope()
```

Seem normal enough.

## Checks on distance perception effects

Here we test if the effects seen in the distance perception are consistent across top and bottom foils, hemifields and eye of origin.

### Foil location

The bottom location for the foil was sometimes a little harder to see, and it has a larger angular difference with the target pair. Does that matter for the results?

```{r, fig.width=10, fig.height=5}
plotSplitDistancePsychometricCurves()
```

Somehow the difference between ipsi- and contralateral presentation goes away when the foil is in the bottom position.

And the statistics:

```{r}
doSplitDistanceStats()
```

PSE shifts depending on target location (across blind spot or elsewhere), and this is modulated by the location of the foil. No effects on slope.

### Hemifield

Here we compare if the left run and right run are different.

```{r, fig.width=10, fig.height=5}
plotHemiDistancePsychometricCurves()
```

Seems that the bias effect of the blind spot might be a little bit smaller in the left hemifield. Let's test this.

```{r}
doHemiDistanceStats()
```

The PSE is still affected by the blind spot, but also by hemifield. The two also interact. No effects on slope.

### Eye of origin

Here we compare the left and right eye of origin. Note that the curves/data here should be the same as in the previous check (plot and stats), just shuffled in how they are assigned to conditions / factors.

```{r fig.width=10, fig.height=5}
plotOrigDistancePsychometricCurves()
```

Looks like the main effect is still there, but might depend a bit on the eye of origin.

```{r}
doOrigDistanceStats()
```

The effect of the location of the target pair persists. While there is no main effect of the eye of origin on the PSE, it does interact with the location of the target. No effects on slope.

# Effect strength across participants

Finally, we do a cursory check to see if participants with a large effect in the distance perception task also have a large effect in the curvature task. We do this by plotting the difference between the two conditions (at blind spot / away from blind spot) in the ipsi-lateral eye only for each participant in both tasks.

```{r fig.width=7, fig.height=5}
plotEffectStrength()
```

That doesn't look like much. Here is a regression on this data:

```{r}
regressionEffectStrength()
```

The NHST regression fails to reject the null hypothesis. The Bayesian linear model does not show evidence for the null hypothesis either.

## Export figures

Each figure can be saved in various formats, for now, we just do PDFs.

```{r}

for (target in c('pdf', 'svg')) {
  plotAreaPsychometricCurve(target=target)
  plotAreaDensityCurve(target=target)
  plotCurvaturePsychometricCurve(target=target)
  plotDistancePsychometricCurve(target=target)
  
  # checks, might not need them:
  plotDistancePSEslope(target=target)
  plotSplitDistancePsychometricCurves(target=target)
  plotHemiDistancePsychometricCurves(target=target)
  plotOrigDistancePsychometricCurves(target=target)
}

```

