---
title: "Intrepid 2a replication"
authors: "Marius 't Hart, Patrick Cavanagh"
date: "2025-04-14"
format: pptx
# format: revealjs
editor: visual
---

## Overview

```{r eval=T, echo=FALSE}
# this script has function to download the data:
source("R/data.R")
# this script contains generic utility functions:
source("R/utilities.R")
# script with custom psychometric functions:
source("R/mprobit.R")

# this script contains functions to process the distance task data:
source('R/distance2.R')
source('R/distance3.R') # checks on original tests
# this script contains functions to process the curvature task data:
source('R/curvature.R')
# this script contains functions to process the area task data:
source("R/area.R")
# alternative just using a Gaussian density distribution:
source("R/area2.R")

# scripts testing relations between effects in distance and curvature task:
source("R/dist_curve.R")

# set up parallelization:
parallelize <- TRUE
if (parallelize) {
  ncores   <- parallel::detectCores()
  usecores <- max(c(1,floor(ncores*0.5)))
  # usecores <- 16
  cluster    <- parallel::makeCluster(usecores)
} else {
  cluster <- NULL
}
```

::: {.incremental}
- Almost complete (1 more participant)
- Fit a probit function to each participant's data
- Allows separate tests on bias and slope
- Custom fitting for the distance task
:::

## Area perception

:::: {.columns}

::: {.column width="67%" .incremental}
```{r fig.width=6, fig.height=5}
plotAreaPsychometricCurve()
```
:::

::: {.column width="33%" .incremental}
- no effects
:::

::::

## Curvature perception data

:::: {.columns}

::: {.column width="67%" .incremental}
```{r fig.width=6, fig.height=5}
plotCurvaturePsychometricCurve()
```
:::

::: {.column width="33%" .incremental}
- small effect of location (away vs across) on bias
- no effects on slope
:::

::::

## Distance perception data

:::: {.columns}

::: {.column width="67%" .incremental}
```{r fig.width=6, fig.height=5}
plotDistancePsychometricCurve(cluster=cluster)
```
:::

::: {.column width="33%" .incremental}
- effect of location (away vs across) on bias
- effect of eye (ipsi vs contra) on bias at the across locations
- small effect of location on slope
:::

::::

## Thanks!


## Distance parameter distributions

:::: {.columns}

::: {.column width="75%" .incremental}
```{r fig.width=8, fig.height=5}
plotDistancePSEslope()
```
:::

::: {.column width="25%" .incremental}
 
:::

::::

## Distance effect of foil location

:::: {.columns}

::: {.column width="75%" .incremental}
```{r fig.width=9, fig.height=5}
plotSplitDistancePsychometricCurves()
```
:::

::: {.column width="25%" .incremental}
- foil location (above / below blind spot) has some effect
- main effect of location remains
:::

::::

## Distance effect of stimulated eye

:::: {.columns}

::: {.column width="75%" .incremental}
```{r fig.width=9, fig.height=5}
plotOrigDistancePsychometricCurves()
```
:::

::: {.column width="25%" .incremental}
- stimulated eye relates to filter used (green vs. red)
- main effect of location remains
:::

::::

## Distance effect of hemifield

:::: {.columns}

::: {.column width="75%" .incremental}
```{r fig.width=9, fig.height=5}
plotHemiDistancePsychometricCurves()
```
:::

::: {.column width="25%" .incremental}
- essentially the same as splitting by eye of origin
- main effect of location remains
:::

::::

## Distance effect of first dot-pair

:::: {.columns}

::: {.column width="75%" .incremental}
```{r fig.width=9, fig.height=5}
plotFirstDistancePsychometricCurves()
```
:::

::: {.column width="25%" .incremental}
- some small differences, but
- main effect of location remains
:::

::::

## Area perception bias

```{r}
doAreaBiasANOVA()
```

## Area perception slope

```{r}
doAreaSlopeANOVA()
```

## Curvature perception bias

```{r}
doCurvatureBiasANOVA()
```

## Curvature perception slope

```{r}
doCurvatureSlopeANOVA()
```

## Distance perception bias

```{r}
doDistanceBiasANOVA()
```

## Distance perception bias follow-up

```{r}
doDistanceBiasANOVAfollowup()
```

## Distance perception slope

```{r}
doDistanceSlopeANOVA()
```

## Summary of findings

::: {.incremental}
- Area:
  - no significant effects
- Curvature:
  - small effect of location (away vs across) on bias
- Distance
  - significant effect of location (ipsi vs contra) on bias
  - contralateral eye shows larger effect than ipsilateral
  - small effect of location on slope
:::


